Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-14 paper/intro.tex
Wordsmiting; added CakeML infinitary proof papers

In recent years, aerospace stakeholders have become aware that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  In addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}. Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.  Cyber vulnerabilities are often discovered during penetration testing late in the development process.  Worse yet, they may be discovered after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
  views expressed are those of the authors and do not reflect the
  official policy or position of the Defense Advanced
  Research Projects Agency (DARPA) or the U.S. Government.}
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to satisfy requirements added to mitigate cyber vulnerabilities. In this work, the behavior of the system is modeled by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis with the new requirements for mitigation.  The requirements assume properties on the input and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement such requirements, AGREE fails to prove the system meets these new requirements.

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, which is a verified implementation of ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

@@

\begin{comment}
Here is the basic process:
\begin{enumerate}
  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
  \item Contiguity types include dependent types and the ability to define when a type is malformed
  \item Define a contract language that makes assumptions on input and guarantees properties on output
  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
  \item Show how the contract can be expressed in terms of a step function 
  \item The step-function is lifted to traces
  \item A contract defines a language over traces
  \item A step function recognizes membership in that language (prefix closed)
  \item Follow the same pattern used in contiguity types
  \item The CakeML comes for free once the HOL4 theories are shown
\end{enumerate}

>>>>

In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}. Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.  Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) program is focused on developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
  views expressed are those of the authors and do not reflect the
  official policy or position of the Defense Advanced
  Research Projects Agency (DARPA) or the U.S. Government.}
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE}, based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  The AGREE analysis \emph{assumes}  properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.  Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

%% DSH  This paragraph seems mostly duplicative of the paragraph above; attempted to merge the two.
%A novel aspect of the BriefCASE approach is the ability to automatically transform the model to insert high-assurance %components to satisfy requirements added to mitigate cyber vulnerabilities. In this work, the behavior of the system is %modeled by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result %of the cyber-vulnerability analysis with the new requirements for mitigation.  The requirements assume properties on the %input and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to %implement such requirements, AGREE fails to prove the system meets these new requirements.

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

\begin{comment}
The basic process proceeds as follows:
\begin{enumerate}
  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
  \item Contiguity types include dependent types and the ability to define when a type is malformed
  \item Define a contract language that makes assumptions on input and guarantees properties on output
  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
  \item Show how the contract can be expressed in terms of a step function 
  \item The step-function is lifted to traces
  \item A contract defines a language over traces
  \item A step function recognizes membership in that language (prefix closed)
  \item Follow the same pattern used in contiguity types
  \item The CakeML comes for free once the HOL4 theories are shown
\end{enumerate}
